---
title: "Methods for multimodality in size dists"
output: html_notebook
---

```{r setup}
portal_data <- neonbecs::get_toy_portal_data()

library(dplyr)
library(ggplot2)
```

## Make "ordinary" and "weighted" size dists

```{r make size dists}
osd <- portal_data %>%
  group_by(individual_species_ids) %>%
  summarise(mean_size = mean(individual_sizes)) %>%
  ungroup()

wsd <- portal_data %>%
  group_by(individual_species_ids) %>%
  summarise(mean_size = mean(individual_sizes),
            abundance = n()) %>%
  ungroup()
```

```{r plot size dists}
osd_plot <- ggplot(data = osd, aes(x = mean_size)) +
  geom_histogram(data = osd, inherit.aes = TRUE) + 
  theme_bw()


osd_plot

```

## KS test for uniformity

```{r ks test}
ks_uniform <- ks.test(osd$mean_size, y = punif)
print(ks_uniform)
```


RMD is curious how draws from a uniform distribution would stack up here

```{r ks bootstrap}
ks_uniform2 <- ks.test(runif(n = 7, 
                             min = 5,
                             max = 150), 
                       y = punif)
print(ks_uniform2)
```
Draws from a bounded uniform distribution come up as significantly different from `punif` in `ks.test`. 

## diptest

```{r diptest}
library(diptest)

dip_osd <- dip.test(osd$mean_size,  
                    simulate.p.value = TRUE,
                    B = 2000)
print(dip_osd)

dip_osd2 <- dip.test(runif(runif(n = 7, 
                             min = 5,
                             max = 150)),
                     simulate.p.value = TRUE,
                     B = 2000)
print(dip_osd2)
```

So based on this we would be trying to fit a multimodal distribution *to either the real data or ```runif()```!*

## KDE

```{r kde}

kde_osd <- density(osd$mean_size, bw = "nrd0", kernel = c("gaussian"))
plot(kde_osd)

library(pastecs)

kde_osd_forturnpoints <- kde_osd$y[seq(0, 500, by = 5)]

kde_osd_points <- turnpoints(kde_osd_forturnpoints)
summary(kde_osd_points)
```

KDE method finds 2 peaks in real data.
```{r kde uniform}

kde_uniform <- density(runif(n = 7,
                             min = 5,
                             max = 150), bw = "nrd0", kernel = c("gaussian"))
plot(kde_uniform)

kde_uniform_forturnpoints <- kde_uniform$y[seq(0, 500, by = 5)]

kde_uniform_points <- turnpoints(kde_uniform_forturnpoints)
summary(kde_uniform_points)
```

## Latent class analysis
Called Latent Profile Analysis for non-binary data. (RMD thinks.)

```{r lca}
library(tidyLPA)
lpa_osd <- estimate_profiles(osd$mean_size, 4)
```

